Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-01-08 manuscript/manuscript.Rmd
updated intro

Introduction
------------
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Clark2010; @Freckleton2011; @Queenborough2011; @Clark2012]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations [@Clark2004; @Freckleton2011]. Such population-level data will never match the precision of individual-level data, but it is more feasible to attain a broad coverage sample when collecting coarse-scale data. This presents a difficult trade-off: on the one hand, individual-level data leads to more reliable models; on the other hand, population-level data leads to models that will produce less precise predicitons but can be applied over greater spatial and temporal extents. If the goal of the model is to predict the consequences of climate change, then we argue that in the face of limited time and money population-level data are the best option. However, the performance of models based on population-level data have rarely been compared to that of models based on individual-level data.

@@

Here, we test the hypothesis (idea?) that statistical and population models based on aggregated data can detect climate signals as wells as models based on individual-level data. We use a unique demographic dataset that tracks the fates of individual plants from four species to build single-species population models, since those are often used tools for ecological forecasts and climate vulnerability assessments. We first fit population models with interannual variation in vital rates explained, in part, by climate covariates. We then perturb the climate covariates to test the sensitivities of species to climate change. By doing these analyses using both individual and aggregated forms of the same data, we provide a rigorous test of our hypothesis. We find that...

>>>>

Introduction
------------
Population models are important tools for predicting the impacts of environmental change on species. But reconciling the scales at which population models are parameterized and the scales at which environmental changes play out remains a challenge [@Clark2010; @Freckleton2011; @Queenborough2011; @Clark2012]. The major hurdle is that most population models, at least for plant species, are built using data from small, localized plots because parameterizing traditional population models requires tracking the fates of individuals. These models are difficult to scale up from the micro to meso-scales because the fitted parameters do not fully represent the spatial variation present at scales beyond that at which the data are collected [@Sæther2007]. At the same time, most demographic data is collected over short time spans. For example, the most common study duration in the COMPADRE matrix population model database is 4 years and only a few exceed 10 years [@Salguero-Gomez2015]. The constrained spatio-temporal extent of most demographic datasets reflects the difficulty of collecting such data, but those constraints limit our ability to extrapolate population models. Thus, our ability to use population models to predict the consequences of climate change is limited when we rely on individual-level data.

Aggregate measures of individual plant performance, such as those typically collected as part of large-scale census efforts, offer an alternative to detailed demographic data for modeling populations [@Clark2004; @Freckleton2011]. Such population-level data will never match the precision of individual-level data, but it is more feasible to attain a broad coverage sample when collecting coarse-scale data. This presents a difficult trade-off: on the one hand, individual-level data leads to more reliable models; on the other hand, population-level data leads to models that will produce less precise predicitons but can be applied over greater spatial and temporal extents. An open question is how well models based on population-level data compare to models based on individual-level data.

@@

Here, we test the assumption that statistical and population models based on aggregated data can detect climate signals as wells as models based on individual-level data. We use a unique demographic dataset that tracks the fates of individual plants from four species over 14 years to build single-species population models, since those are often used tools for ecological forecasts and climate vulnerability assessments. We first fit population models with interannual variation in vital rates explained, in part, by climate covariates. We then perturb the climate covariates to test the sensitivities of species to climate change. By doing these analyses using both individual and aggregated forms of the same data, we provide a rigorous test of our hypothesis. We find that...

### 2015-01-08 manuscript/manuscript.tex
updated intro

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The major
hurdle is that most population models, at least for plant species, are
built using data from small, localized plots because parameterizing
traditional population models requires tracking the fates of
individuals. These models are difficult to scale up from the micro to
meso-scales because the fitted parameters do not fully represent the
spatial variation present at scales beyond that at which the data are
collected (Sæther et al. 2007). Thus, our ability to use population
models to predict the consequences of climate change is limited when we
rely on individual-level data.

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations (Clark
and Bjørnstad 2004, Freckleton et al. 2011). Such population-level data
will never match the precision of individual-level data, but it is more
feasible to attain a broad coverage sample when collecting coarse-scale
data. This presents a difficult trade-off: on the one hand,
individual-level data leads to more reliable models; on the other hand,
population-level data leads to models that will produce less precise
predicitons but can be applied over greater spatial and temporal
extents. If the goal of the model is to predict the consequences of
climate change, then we argue that in the face of limited time and money
population-level data are the best option. However, the performance of
models based on population-level data have rarely been compared to that
of models based on individual-level data.

@@

Here, we test the hypothesis (idea?) that statistical and population
models based on aggregated data can detect climate signals as wells as
models based on individual-level data. We use a unique demographic
dataset that tracks the fates of individual plants from four species to
build single-species population models, since those are often used tools
for ecological forecasts and climate vulnerability assessments. We first
fit population models with interannual variation in vital rates
explained, in part, by climate covariates. We then perturb the climate
covariates to test the sensitivities of species to climate change. By
doing these analyses using both individual and aggregated forms of the
same data, we provide a rigorous test of our hypothesis. We find
that\ldots{}

@@

All of our analyses (model fitting and simulating) were conducted in R
(R Development Core Team 2013). We used the MCMC sampler in JAGS
(Plummer 2003) to estimate the posterior distributions of model
parameters and the package `r2jags' (Su and Yajima 2012) to connect R to
JAGS. We obtained posterior distributions for all model parameters from
three parallel MCMC chains run for 50,000 iterations, after discarding
an initial 50,000 iterations. We assessed convergence visually and using
the Gelman and Rubin (1992) diagnostic in the R package `coda' (Plummer
et al. 2006). Scale reduction factors for all parameters were less than
1.02, indicating convergence. For the purposes of introducing
stochasticity in our population models, we saved the final 1,000
iterations from each chain for all parameters to be used as randomly
drawn values during population simulation.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Boxplots of model residuals for one-step-ahead forecasts at
each observation year. Each one-step forecast was simulated
\texttt{r nSims} times. Note that the y-axes vary across panels. The
light blue line shows the difference between the observed-year percent
cover and the average cover observed across all years. The models tend
to underpredict and perform poorly when observed cover in a given year
is a large deviant from the mean.}
\end{figure}

@@

R Development Core Team. 2013. R Development Core Team. R: A Language
and Environment for Statistical Computing.

@@

>>>>

Population models are important tools for predicting the impacts of
environmental change on species. But reconciling the scales at which
population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The major
hurdle is that most population models, at least for plant species, are
built using data from small, localized plots because parameterizing
traditional population models requires tracking the fates of
individuals. These models are difficult to scale up from the micro to
meso-scales because the fitted parameters do not fully represent the
spatial variation present at scales beyond that at which the data are
collected (Sæther et al. 2007). At the same time, most demographic data
is collected over short time spans. For example, the most common study
duration in the COMPADRE matrix population model database is 4 years and
only a few exceed 10 years (Salguero-Gómez et al. 2015). The constrained
spatio-temporal extent of most demographic datasets reflects the
difficulty of collecting such data, but those constraints limit our
ability to extrapolate population models. Thus, our ability to use
population models to predict the consequences of climate change is
limited when we rely on individual-level data.

Aggregate measures of individual plant performance, such as those
typically collected as part of large-scale census efforts, offer an
alternative to detailed demographic data for modeling populations (Clark
and Bjørnstad 2004, Freckleton et al. 2011). Such population-level data
will never match the precision of individual-level data, but it is more
feasible to attain a broad coverage sample when collecting coarse-scale
data. This presents a difficult trade-off: on the one hand,
individual-level data leads to more reliable models; on the other hand,
population-level data leads to models that will produce less precise
predicitons but can be applied over greater spatial and temporal
extents. An open question is how well models based on population-level
data compare to models based on individual-level data.

@@

Here, we test the assumption that statistical and population models
based on aggregated data can detect climate signals as wells as models
based on individual-level data. We use a unique demographic dataset that
tracks the fates of individual plants from four species over 14 years to
build single-species population models, since those are often used tools
for ecological forecasts and climate vulnerability assessments. We first
fit population models with interannual variation in vital rates
explained, in part, by climate covariates. We then perturb the climate
covariates to test the sensitivities of species to climate change. By
doing these analyses using both individual and aggregated forms of the
same data, we provide a rigorous test of our hypothesis. We find
that\ldots{}

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.02,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each
chain for all parameters to be used as randomly drawn values during
population simulation.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_1.pdf}
\caption{Boxplots of model residuals for one-step-ahead forecasts at
each observation year. Each one-step forecast was simulated
\texttt{r nSims} times. Note that the y-axes vary across panels.}
\end{figure}

@@

Salguero-Gómez, R., O. R. Jones, C. R. Archer, Y. M. Buckley, J.
Che-Castaldo, H. Caswell, D. Hodgson, A. Scheuerlein, D. A. Conde, E.
Brinks, H. de Buhr, C. Farack, F. Gottschalk, A. Hartmann, A. Henning,
G. Hoppe, G. Römer, J. Runge, T. Ruoff, J. Wille, S. Zeh, R. Davison, D.
Vieregg, A. Baudisch, R. Altwegg, F. Colchero, M. Dong, H. de Kroon,
J.-D. Lebreton, C. J. E. Metcalf, M. M. Neel, I. M. Parker, T. Takada,
T. Valverde, L. A. Vélez-Espino, G. M. Wardle, M. Franco, and J. W.
Vaupel. 2015. The compadrePlant Matrix Database: an open online
repository for plant demography. Journal of Ecology 103:202--218.

@@

Team, R. 2013. R Development Core Team. R: A Language and Environment
for Statistical Computing.
